---
title: "315HW8"
author: "Elan Satchit"
date: "2025-04-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(boot)
```

[GITHUB REPO](https://github.com/elansatchit/315HW8)

# Problem 1

## Part A
```{r}
creatinine <- read.csv("creatinine.csv")
model <- lm(creatclear ~ age, data = creatinine)

predict(model, newdata = data.frame(age = 55))
```
We use the fitted linear regression model:

creatclear =b0+b1 x age

Using the estimated coefficients from the model, we plug in age = 55:

## Part B
```{r}
coef(model)[2]
```
The slope of the regression line (coefficient on age) represents the change in clearance per year of age. This tells us that creatinine clearance decreases by approximately 0.62 mL/min per year of age.

## Part C
```{r}
pred_40 <- predict(model, newdata = data.frame(age = 40))
pred_60 <- predict(model, newdata = data.frame(age = 60))
res_40 <- 135 - pred_40
res_60 <- 112 - pred_60
if (res_40 > res_60) "40-year-old is healthier" else "60-year-old is healthier"
```
We predict expected values for both ages using the model and compute residuals (actual - predicted). The person with the larger positive residual is healthier relative to their expected value:

if (res_40 > res_60) "40-year-old is healthier" else "60-year-old is healthier"

This statement prints the correctly modeled statement

# Problem 2

## Part 1 - Italy
```{r}
covid <- read.csv("covid.csv")
italy <- covid %>% filter(country == "Italy")

model_italy <- lm(log(deaths) ~ days_since_first_death, data = italy)
growth_italy <- coef(model_italy)[2]
doubling_italy <- log(2) / growth_italy

boot_growth_italy <- numeric(1000)

for (i in 1:1000) {
  sample_rows <- sample(nrow(italy), replace = TRUE)
  sample_data <- italy[sample_rows, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = sample_data)
  boot_growth_italy[i] <- coef(model)[2]
}

ci_growth_italy <- quantile(boot_growth_italy, c(0.025, 0.975))
ci_doubling_italy <- log(2) / ci_growth_italy

list(
  growth_rate = round(growth_italy, 3),
  growth_ci = round(ci_growth_italy, 3),
  doubling_time = round(doubling_italy, 1),
  doubling_ci = round(ci_doubling_italy, 1)
)
```
We estimate the growth rate of COVID-19 deaths in Italy by fitting a linear model to the log-transformed daily deaths. We bootstrap by manually resampling the data 1000 times, refitting the model each time, and collecting the slope. The 95% confidence interval for the growth rate is taken from the 2.5th and 97.5th percentiles of the bootstrap distribution. 

## Part 2 - Spain
```{r}
spain <- covid %>% filter(country == "Spain")
model_spain <- lm(log(deaths) ~ days_since_first_death, data = spain)
growth_spain <- coef(model_spain)[2]
doubling_spain <- log(2) / growth_spain

boot_growth_spain <- numeric(1000)

for (i in 1:1000) {
  sample_rows <- sample(nrow(spain), replace = TRUE)
  sample_data <- spain[sample_rows, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = sample_data)
  boot_growth_spain[i] <- coef(model)[2]
}

ci_growth_spain <- quantile(boot_growth_spain, c(0.025, 0.975))
ci_doubling_spain <- log(2) / ci_growth_spain

list(
  growth_rate = round(growth_spain, 3),
  growth_ci = round(ci_growth_spain, 3),
  doubling_time = round(doubling_spain, 1),
  doubling_ci = round(ci_doubling_spain, 1)
)
```
The same bootstrapping process is used for Spain. Each iteration involves resampling the dataset with replacement, refitting the model, and storing the slope. This gives us a distribution of growth rate estimates from which we compute confidence intervals.

## Part 3
```{r}
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(title = "COVID-19 Daily Deaths by Country",
       x = "Days Since First Death",
       y = "Reported Deaths") +
  theme_minimal()
```

# Problem 3
```{r}
milk <- read.csv("milk.csv")
milk <- milk %>%
  mutate(price_elasticity = log(price), log_sales = log(sales))

model <- lm(log_sales ~ price_elasticity, data = milk)
elasticity <- coef(model)[2]


boot_elasticity <- numeric(1000)
for (i in 1:1000) {
  sample_rows <- sample(nrow(milk), replace = TRUE)
  sample_data <- milk[sample_rows, ]
  model_boot <- lm(log_sales ~ price_elasticity, data = sample_data)
  boot_elasticity[i] <- coef(model_boot)[2]
}
ci_elasticity <- quantile(boot_elasticity, c(0.025, 0.975))

list(
  elasticity = round(elasticity, 3),
  ci = round(ci_elasticity, 3)
)
```
To estimate the price elasticity of demand for milk, I fit a model that relates the quantity of milk purchased to its price using a power-law relationship: 

Q = KP^b, where b is the elasticity

by fitting this model to the data. To measure uncertainty, I performed bootstrapping by repeatedly sampling the data with replacement 1000 times and refitting the model each time to get a distribution of elasticity estimates.
